---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# verhoeff

This package is my take on implementing the Verhoeff algorithim for calculating check digits. 
My influence for the algorithims implementation was primarily from the wikipedia outline (https://en.wikipedia.org/wiki/Verhoeff_algorithm), and the paper by A. Dunn (Computer program for the calculation and validation of Verhoeff check digits, https://www.ccamlr.org/en/wg-fsa-sam-05/11)

Verhoeff is still undergoing most of its development. A CRAN release will be prepared after I've burned through most of the GH issues

## Installation

`verhoeff` is not (yet) on CRAN. You can install verhoeff from github with:

```{r gh-installation, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
devtools::install_github("condwanaland/verhoeff")
```

## Example

To calculate the Verhoeff Check Digit for a given number, use `verhoeff_calculate`

```{r example}
library(verhoeff)
verhoeff::verhoeff_calculate(123)
```

To calculate for many numbers, just pass a vector.

```{r}
verhoeff_calculate(c(1234, 5678, 9, 10))
```

If you would prefer the output to be in a list, with one check digit per element, use the `as_list` parameter.

```{r}
verhoeff_calculate(c(1234, 5678, 9, 10), as_list = TRUE)
```


The output of `verhoeff_calculate` is designed so it can be a new column in a dataframe (i.e., as the output of a `dplyr::mutate` call)
```{r}
suppressPackageStartupMessages(library(dplyr))

# Make a random dataframe that has integer columns
mtcars$name <- rownames(mtcars)
mtcars %>% 
  select(name, gear) %>% 
  mutate(check_digit = verhoeff_calculate(gear))
```


## Other implementations

See https://github.com/fascinatingfingers/CheckDigit for another implementation of check digits that considers more than just the Verhoeff algorithim.

